---
title: "Homework 5"
date: 'Due: Nov 30, 2:00pm'
output:
 pdf_document:
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1 (18 regular pts)

Single tree classifier on handwritten digit data. Since we are working on a classification problem, please set `method="class"` for the `rpart` function. 

**(a) (6 pts)** Load `zip.RData` into `R` and train a classification tree using training data. (You need to set a small `cp` parameter, say `cp=0.0001`, for the `rpart` function.) 

**Answer:** 

```{r, echo=TRUE}
library('rpart')
set.seed(10)
load("~/Downloads/zip.RData")

tree <- rpart(V1 ~., data = dat$train, method = 'class', cp = 0.0001)
```

**(b) (6 pts)** Use function `printcp` to find the `cp` value that gives the smallest `xerror`.

**Answer:** 

```{r, echo=TRUE}
printcp(tree)
```
cp = 0.00077101 gives xerror of 0.053200.

**(c) (6 pts)** Prune the tree with the `cp` found in part (b). Plot the pruned tree.

**Answer:** 

```{r, echo=TRUE}
tree_p <- prune(tree, cp = 0.00077101)
plot(tree_p, main = "Pruned decision Tree")
text(tree_p)
```

## Question 2 (18 regular pts)

Random Forest on handwritten digit data.

**(a) (6 pts)** Applying random forest on the training data with `ntree=1000`.

**Answer:** 

```{r, echo=TRUE}
library("randomForest")
set.seed(10)
model_rf <- randomForest(V1 ~., data=dat$train, ntree=1000)
```

**(b) (6 pts)** How many number of trees are aggregated to give the lowest OOB error? What is the value of lowest OOB error?

**Answer:** 

```{r, echo=TRUE}
ntree.best <- which.min(model_rf$err.rate[, 'OOB'])
ntree.best
```

**(c) (6 pts)** Re-train your random forest using the optimum number of trees you got in Part (b). Report the misclassification rate on the test data.

**Answer:** 

```{r, echo=TRUE}
model_rf1 <- randomForest(V1 ~., data=dat$train, ntree=262)
y_test <- predict(model_rf1, newdata = dat$test)
misclass <- sum(y_test!=dat$test[,1])/length(y_test)
misclass
```


## Question 3 (12 regular pts)
$k$-means clustering on handwritten digit data. We will only use the training data for this problem. 

**(a) (6 pts)** Use $k$-means clustering with `centers = 3` and `nstart=20` to cluster the data.

**Answer:** 

```{r, echo=TRUE}
set.seed(10)
res.kmeans <- kmeans(dat$train, centers=3, nstart = 20)
```

**(b) (6 pts)** Plot the centroids found in part (a) as images (there are 3 centroids in total, each of length 256). What digits do they look like?

**Answer:** 

```{r, echo=TRUE}
# recall that we use the following function to plot images
centers <- res.kmeans$centers

conv.image <- function(vec)
{
	mat <- matrix(as.numeric(vec), nrow=16, ncol=16)
	mat <- -mat[, 16 : 1]
	image(mat, col=gray(seq(0,1, 0.01)), xaxt='n', yaxt='n')
}
par(mar=c(0, 0, 0, 0), mfrow=c(1,3))

## Your code
for(i in range(3)){
  conv.image(centers[i,])
}
```


#They look like the number 4.







